Comparing Reinforcement Learning and Finite State Machine Agents in Real Time Strategy Games: Impact on Player Experience

### 2025-04-18 Paper/main.tex
research methodology needs a little more work, and to use the notes about the literature and convert it into a properly structured literature review.

@@

% TODO:

@@

Create a simple RTS game, and implement both RL and FSM agents. Conduct a deductive experiment, where players will be split into 2 groups,
where the first group will play against the FSM, fill out a questionnaire to gather data on their experience, and then play against the RL agent,
and again answer the questionnaire. The second group will do the same, but in reverse order, playing against the RL agent first, and then the FSM.
The end of each experiment's questionnaires will have an open ended question, where players can provide feedback on their experience,
which will be used to gather qualitative data on their experience. Qualitative data will be used to support the quantitative data gathered from
the questionnaires, as well as some in-game analytics and logs. This should help capture the more qualitative aspects of player experience,
while still keeping everything quantitative and measurable. This should help in answering the research questions, and in determining the validity
of the Hypothesis.

@@

% TODO:

>>>>

% Have notes about papers locally, but have not formulated them into a proper literature review yet. Should focus on discussing their results, as most fall between the above 2 mentioned papers, and another one
% which takes a different approach, which might be out of scope. Discuss the advantages and disadvantages of each approach, being balancing the difficulty using snapshots, and building DDA into the agent's
% reward structure (fighting game paper)

@@

This study adopts a **positivist research philosophy**, which means it emphasizes the use of objective measurements and observable phenomena, and it is suitable for this study as it aims
to evaluate the impact of RL and FSM AI opponents on player experience through measurable data, aligning it with **quantitative methodology**. It follows a **deductive approach**, with an **experimental research design**,
as it starts with a hypothesis, and then using an experimental prototype, tests the hypothesis by allowing participants to play it, and then uses the data to observe the difference the independent variable (AI type)
has on the dependent variables (player experience, immersion, and perceived difficulty). This will then be used to answer the research questions and prove the hypothesis. It is important to note that the study
will also gather a very small amount of qualitative data, making it a **mixed-methods approach**. This is done to help align the more human aspects of player experience with the more quantitative aspects.

@@

% Is more to be done?
Create a simple RTS game, and implement both RL and FSM agents. Conduct a deductive experiment, where players will be split into 2 groups,
where the first group will play against the FSM, fill out a questionnaire to gather data on their experience, and then play against the RL agent,
and again answer the questionnaire. The second group will do the same, but in reverse order, playing against the RL agent first, and then the FSM.
The end of each experiment's questionnaires will have an open ended question, where players can provide feedback on their experience,
which will be used to gather qualitative data on their experience. Qualitative data will be used to support the quantitative data gathered from
the questionnaires, as well as some in-game analytics and logs. This should help capture the more qualitative aspects of player experience,
while still keeping everything quantitative and measurable. This should help in answering the research questions, and in determining the validity
of the Hypothesis.

@@

**Validity**: The study ensures validity by desigining the experiment to control for confounding variables, such as the order in which participants play, since on the second playtest they would be more
familiar with the game mechanics and controls. The use of validating the survey results with in-game analytics and engagement metrics further streghtens the validity of the research being conducted.

**Reliability**: To help ensure reliability, the same setup, game environment and surveys will be used for all participants. Any and all instructions will be standardised, using text and/or video/audio
recordings to ensure that all participants are given the same instructions, without any bias that comes from the researcher. All datacollection processes will be automated, further standardising the process
and ensuring reliability.

**Generalizability/Transferability**: While the finding will be focused on the specific RTS game developed for this study, the insights gained can be generalised to other games within the RTS genre,
which follow similar game mechanics. The results may also be applicable to other game genres with similar AI opponent implementations, such as turn-based strategy games, but the transferability
of results could be limited, and as such should be explored in future research.
